---
title: "p8105_hw6_xl3141"
author: "Xinyuan Liu"
date: "11/28/2021"
output: github_document
---

## Set up

```{r}
library(tidyverse)
library(modelr)
library(mgcv)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

```{r}
birthweight_df = read_csv("birthweight.csv")

birthweight_df= 
  birthweight_df %>% 
    drop_na() %>% 
    mutate(
      babysex = fct_recode(factor(babysex), male = "1", female = "2"),
      frace = fct_recode(factor(frace), White = "1", Black = "2", Asian = "3", Puerto_Rican = "4", Other = "8"),
      malform = fct_recode(factor(malform), absent = "0", present = "1"),
      mrace = fct_recode(factor(mrace), White = "1", Black = "2", Asian = "3", Puerto_Rican = "4")
    )
```

From my point of view, the child's birthweight should be associated with baby's sex, baby's head circumference at birth, baby's length at birth, gestational age, mother's age at menarche, mother's race, father's race, and mother's weight gain during pregnancy. 

First, let's explore if the categorical variables I listed above have some influence on the child's birth weight!

```{r}
birthweight_df %>% 
  select(bwt, babysex, frace, mrace) %>% 
  pivot_longer(
    cols = -bwt,
    names_to = "category",
    values_to = "levels"
  ) %>% 
  ggplot(aes(y = bwt, x = levels)) +
  geom_boxplot() +
  facet_wrap(. ~ category, scale = "free") +
  labs(
    y = "birth weight",
    title = "distribution of birthweight according to different categorical variables"
  )
```

It seems that all of them will affect child's birthweight. 
Then, let's take a look at continuous variables

```{r}
birthweight_df %>% 
  select(bwt, bhead, blength, gaweeks, menarche, wtgain) %>% 
  pivot_longer(
    cols = -bwt, 
    names_to = "variables",
    values_to = "values"
  ) %>% 
  ggplot(aes(x = values, y = bwt)) +
  geom_point(alpha = .3) +
  facet_wrap(. ~ variables, scales = "free")+
  labs(
    y = "birth weight",
    title = "distribution of birthweight according to different continuous variables"
  )
```

It seems that mother's age at menarche does not have a significant impact on child's birthweight from the scatterplot, so I will remove that variable in my model.
 
Here is my model:

```{r}
my_model = lm(bwt ~ bhead + blength + gaweeks + wtgain + babysex + frace + mrace, data = birthweight_df)
```

Let's look at the residuals in our model

```{r}
birthweight_df %>% 
  modelr::add_residuals(my_model) %>% 
  modelr::add_predictions(my_model) %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = .3) +
  labs(
    title = "scatterplot for residuals against fitted values",
    x = "predicted birthweight",
    y = "residual"
  )
```

compare with the other 2 model provided

```{r}
model1 = lm(bwt ~ blength + gaweeks, data = birthweight_df)
model2 = lm(bwt ~ bhead + babysex + blength + bhead * babysex + bhead * blength + babysex * blength + babysex * blength * bhead, data = birthweight_df)
```

Test the model fit

```{r}
set.seed(2)
cv_df = 
  crossv_mc(birthweight_df, 100) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )

cv_df = 
  cv_df %>% 
  mutate(
    my_model = map(.x = train, ~ lm(bwt ~ bhead + blength + gaweeks + wtgain + babysex + frace + mrace, data = .x)),
    model1 = map(.x = train, ~lm(bwt ~ blength + gaweeks, data = .x)),
    model2 = map(.x = train, ~lm(bwt ~ bhead + babysex + blength + bhead * babysex + bhead * blength + babysex * blength + babysex * blength * bhead, data = .x))
    )%>% 
  mutate(
    rmse_my_model = map2_dbl(.x = my_model, .y = test, ~rmse(model = .x, data = .y)),
    rmse_model1 = map2_dbl(.x = model1, .y = test, ~rmse(model = .x, data = .y)),
    rmse_model2 = map2_dbl(.x = model2, .y = test, ~rmse(model = .x, data = .y))
  )

cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  ggplot(aes(x = model, y = rmse)) +
  geom_boxplot()
  

```

The plot above compared the prediction error of three models. The model using baby's length at birth has the highest average rmse compared to other two models, indicating a worse fitting of this model. My model seems to have the smallest rmse among the three models which means it predicts the child's birth weight better than other two models provided.

## Problem 2

loading data

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

bootstrap

```{r}
set.seed(3)
bootstrap = 
  weather_df %>% 
  bootstrap(n = 5000, id = "strap_number") 

r_square_df = 
  bootstrap %>% 
    mutate(
        models = map(.x = strap, ~lm(tmax ~ tmin, data = .x)),
        glance = map(models, broom::glance)
    ) %>% 
    unnest(glance) %>% 
    select(strap_number, r.squared)

r_square_df %>% 
  ggplot(aes(x = r.squared)) +
  geom_density() +
  labs(title = "distribution of r squared")
```

The distribution looks symmetric but is not perfectly bell-shaped. It centers around `r r_square_df %>% pull(r.squared) %>% mean()` with variance `r r_square_df %>% pull(r.squared) %>% var()`. The $95\%$ confidence interval is between [`r r_square_df %>% pull(r.squared) %>% quantile(c(0.025, 0.975))`].


```{r}
log_df = 
  bootstrap %>% 
  mutate(
    models = map(.x = strap, ~lm(tmax ~ tmin, data = .x)),
    result = map(models, broom::tidy)
  ) %>% 
  unnest(result)  
  
plot_log =  
  log_df %>% 
    group_by(strap_number) %>% 
    mutate(log_beta = log(prod(estimate))) %>% 
    filter(term == "tmin") %>% 
    select(strap_number, log_beta)

plot_log %>% 
    ggplot(aes(x = log_beta)) +
    geom_density() +
    labs(title = "distribution of log(b0*b1)")


```

The distribution looks bell-shaped and symmetric. It centers around `r plot_log %>% pull(log_beta) %>% mean()` with variance about `r plot_log %>% pull(log_beta) %>% var()`. The $95\%$ confidence interval is between [`r plot_log %>% pull(log_beta) %>% quantile(c(0.025, 0.975))`].